---
title: "R HTTP framework benchmarks"
output:
  github_document:
    toc: true
params:
  run_benchmark: false
  results_file: "results/latest.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>")
```

```{r run-benchmark, include=FALSE}
if (isTRUE(params$run_benchmark)) {
  status <- system2("bash", "tools/bench/run.sh")
  if (!identical(status, 0L)) {
    stop("Benchmark command failed with status ", status, call. = FALSE)
  }
}
```

```{r load-results, include=FALSE}
results_file <- params$results_file
if (!file.exists(results_file)) {
  stop(
    "No benchmark results found at ", sQuote(results_file), ".\n",
    "Run `bash tools/bench/run.sh` first, or render with ",
    "`params = list(run_benchmark = TRUE)`. ",
    call. = FALSE
  )
}

bench <- read.csv(results_file, stringsAsFactors = FALSE, check.names = FALSE)
required <- c(
  "run_id", "generated_at", "variant", "label", "route", "kind",
  "requests_per_sec", "latency_avg", "wrk_args", "server_options",
  "host", "cpu", "logical_cpus", "memory", "os", "r_version",
  "package_versions"
)
missing <- setdiff(required, names(bench))
if (length(missing) > 0L) {
  stop("Benchmark results are missing columns: ", paste(missing, collapse = ", "),
       call. = FALSE)
}

variant_order <- unique(bench$variant)
variant_labels <- setNames(bench$label, bench$variant)

format_rps <- function(x) {
  n <- suppressWarnings(as.numeric(x))
  if (is.na(n)) return(x)
  format(round(n), big.mark = ",", scientific = FALSE, trim = TRUE)
}

cell_for <- function(variant, route) {
  row <- bench[bench$variant == variant & bench$route == route, , drop = FALSE]
  if (nrow(row) == 0L) return(NA_character_)
  sprintf("%s rps, %s", format_rps(row$requests_per_sec[[1]]), row$latency_avg[[1]])
}

comparison <- data.frame(Variant = unname(variant_labels[variant_order]), check.names = FALSE)
comparison[["`/ping` JSON"]] <- vapply(variant_order, cell_for, character(1), route = "/ping")
comparison[["`/ping-text` plain text"]] <- vapply(variant_order, cell_for, character(1), route = "/ping-text")

server_options <- unique(bench[c("label", "server_options")])
names(server_options) <- c("Variant", "Server options")
rownames(server_options) <- NULL

server_scripts <- c(
  "drogonR native" = "servers/drogonR-native.R",
  "drogonR plumber-shim" = "servers/drogonR-plumber-shim.R",
  "plumber" = "servers/plumber.R",
  "plumber2" = "servers/plumber2.R",
  "plumber2 mirai async" = "servers/plumber2-mirai.R",
  "RestRserve" = "servers/restrserve.R",
  "nanonext" = "servers/nanonext.R",
  "httpuv" = "servers/httpuv.R"
)
server_scripts <- server_scripts[file.exists(server_scripts)]

meta <- bench[1L, , drop = FALSE]
```

# R HTTP framework benchmarks

This report is generated from machine-readable benchmark output, not by copying table values by hand.

## Machine and benchmark details

```{r metadata, echo=FALSE, results='asis'}
cat("- Results file: `", results_file, "`\n", sep = "")
cat("- Run id: `", meta$run_id, "`\n", sep = "")
cat("- Generated at: ", meta$generated_at, "\n", sep = "")
cat("- `wrk` args: `", meta$wrk_args, "`\n", sep = "")
cat("- Host: ", meta$host, "\n", sep = "")
cat("- OS: ", meta$os, "\n", sep = "")
cat("- CPU: ", meta$cpu, "\n", sep = "")
cat("- Logical CPUs: ", meta$logical_cpus, "\n", sep = "")
cat("- Memory: ", meta$memory, "\n", sep = "")
cat("- R: ", meta$r_version, "\n", sep = "")
cat("- Package versions: ", meta$package_versions, "\n", sep = "")
```

## Workload

- `GET /ping`: JSON object equivalent to `{"ok": true}` where possible.
- `GET /ping-text`: scalar string `"ok"` through the framework's default response handling.

## Results

```{r comparison-table, echo=FALSE}
knitr::kable(comparison, align = c("l", "r", "r"))
```

## Server options

```{r server-options-table, echo=FALSE}
knitr::kable(server_options, align = c("l", "l"))
```

Raw `wrk` output and server logs are written under `results/` by `tools/bench/run.sh`.

## Server code

For transparency, these are the exact server scripts used by the benchmark.

```{r server-code, echo=FALSE, results='asis'}
for (label in names(server_scripts)) {
  path <- server_scripts[[label]]
  cat("### ", label, "\n\n", sep = "")
  cat("Source: `", path, "`\n\n", sep = "")
  cat("```r\n")
  cat(readLines(path, warn = FALSE), sep = "\n")
  cat("\n```\n\n")
}
```
