---
title: "benchmark-r-http-frameworks"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>")
latest_results <- "results/latest.csv"
if (file.exists(latest_results)) {
  bench <- read.csv(latest_results, stringsAsFactors = FALSE, check.names = FALSE)
  variant_order <- unique(bench$variant)
  variant_labels <- setNames(bench$label, bench$variant)

  format_rps <- function(x) {
    n <- suppressWarnings(as.numeric(x))
    if (is.na(n)) return(x)
    format(round(n), big.mark = ",", scientific = FALSE, trim = TRUE)
  }

  cell_for <- function(variant, route) {
    row <- bench[bench$variant == variant & bench$route == route, , drop = FALSE]
    sprintf("%s rps, %s", format_rps(row$requests_per_sec[[1]]), row$latency_avg[[1]])
  }

  comparison <- data.frame(Variant = unname(variant_labels[variant_order]), check.names = FALSE)
  comparison[["`/ping` JSON"]] <- vapply(variant_order, cell_for, character(1), route = "/ping")
  comparison[["`/ping-text` plain text"]] <- vapply(variant_order, cell_for, character(1), route = "/ping-text")
  server_options <- unique(bench[c("label", "server_options")])
  names(server_options) <- c("Variant", "Server options")
  rownames(server_options) <- NULL
  meta <- bench[1L, , drop = FALSE]
}
```

A standalone benchmark suite for comparing R HTTP / REST serving frameworks on the same tiny endpoints.

It is independent of any framework source checkout: it runs installed R packages and records raw `wrk` output plus machine/package metadata.

## Latest results

```{r latest-results, echo=FALSE, results='asis'}
if (exists("comparison")) {
  cat("Results file: `", latest_results, "`  \n", sep = "")
  cat("Run id: `", meta$run_id, "`  \n", sep = "")
  cat("Generated at: ", meta$generated_at, "  \n", sep = "")
  cat("`wrk` args: `", meta$wrk_args, "`  \n", sep = "")
  cat("Machine: ", meta$os, "; ", meta$cpu, "; ", meta$logical_cpus,
      " logical CPUs; ", meta$memory, "  \n", sep = "")
  cat("R: ", meta$r_version, "  \n", sep = "")
  cat("Packages: ", meta$package_versions, "\n\n", sep = "")
  print(knitr::kable(comparison, align = c("l", "r", "r")))
} else {
  cat("No `results/latest.csv` yet. Run `bash tools/bench/run.sh` and re-render `README.Rmd`.\n")
}
```

## Frameworks

Current variants:

- `drogonR-native` — `drogonR::dr_app()` + R handler closures
- `drogonR-plumber-shim` — a `plumber` router served by `drogonR::pr_run()`
- `plumber` — classic `plumber::pr_run()`
- `plumber2` — `plumber2::api_run()` with synchronous handlers
- `plumber2-mirai` — `plumber2::api_run()` with `async = TRUE` and persistent `mirai` daemons
- `RestRserve` — `RestRserve::Application` served by the `Rserve` backend
- `nanonext` — direct `nanonext::http_server()` endpoint handlers

## Workload

All variants expose:

- `GET /ping` — JSON object equivalent to `{"ok": true}` where possible
- `GET /ping-text` — scalar string `"ok"` through the framework's normal response path

Default load:

```bash
wrk -t4 -c50 -d30s http://127.0.0.1:<port>/<route>
```

Override benchmark load and framework concurrency options with environment variables:

```bash
WRK_THREADS=8 WRK_CONNECTIONS=100 WRK_DURATION=60s bash tools/bench/run.sh
```

Server options:

```bash
# drogonR
DROGONR_THREADS=4 DROGONR_WORKERS=1 bash tools/bench/run.sh

# plumber2 async/mirai variant
PLUMBER2_MIRAI_DAEMONS=5 bash tools/bench/run.sh

# RestRserve / Rserve backend on Linux
RESTRSERVE_JIT_LEVEL=0 RESTRSERVE_PRECOMPILE=false bash tools/bench/run.sh
```

RestRserve uses the Rserve HTTP backend. This benchmark targets Linux; Rserve uses forked request processing for concurrent work. `RESTRSERVE_RSERVE_PORT` can be set if you need to pin the internal Rserve/QAP port instead of letting RestRserve choose one.

```{r server-options, echo=FALSE}
if (exists("server_options")) {
  knitr::kable(server_options, align = c("l", "l"))
}
```

## Requirements

System tools:

```bash
sudo apt-get install curl wrk
```

R packages:

```r
install.packages(c("plumber", "plumber2", "mirai", "RestRserve", "nanonext", "rmarkdown", "knitr"))
# plus drogonR from wherever you install it
```

## Run

Run the benchmark:

```bash
bash tools/bench/run.sh
```

Render the README and full report from latest results:

```bash
Rscript -e 'rmarkdown::render("README.Rmd")'
Rscript -e 'rmarkdown::render("benchmarks.Rmd")'
```

Run and render the full report in one command:

```bash
Rscript -e 'rmarkdown::render("benchmarks.Rmd", params = list(run_benchmark = TRUE))'
```

Outputs are written to `results/`:

- raw `wrk` files per framework/route
- server logs
- `comparison-<timestamp>.csv`
- `latest.csv`

`README.md` is generated from `README.Rmd`. `benchmarks.md` is generated from `benchmarks.Rmd` and includes the server code used for the run.
